---
title: "STAT-656-Project"
author: "Isaac Ke"
date: "4/7/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(656)
```

## Load Packages
```{r, message=FALSE}
require(caret)
require(dplyr)
require(ggplot2)
require(caret)
require(e1071)
require(earth)        # fit MARS models
require(vip)          # variable importance
require(glmnet)
require(lattice)   
require(pROC)
require(corrplot)
require(smotefamily)  # SMOTE
```

## Read Data
```{r}
data = read.csv("~/GitHub/Wine-Quality-Prediction/WineQT.csv")
str(data)
head(data)
```

## Data Structure
``` {r, warning=FALSE}
data$quality = as.factor(data$quality)
data = dplyr::select(data, -Id)
str(data)
```

```{r}
Y_presmote = make.names(data$quality)
X_presmote = select(data, -quality) 
```

# SMOTE

> https://rdrr.io/cran/smotefamily/man/
> https://medium.com/analytics-vidhya/smote-technique-for-unbalanced-data-of-3-classes-in-r-programming-cardiotocography-data-set-474bb5dbf8dd 

## SMOTE round 1 (X3)
Note: If splitting two times (one on Ytrain and one on Ytest), need to change SMOTE() function K parameter to 1 since the test set only has 1 nearest neighbor for the minority class. 
```{r}
smote_data1 = SMOTE(X_presmote, Y_presmote, K=2, dup_size = 80)$data
smote_data1 = rename(smote_data1, quality = class)
smote_data1$quality = as.factor(smote_data1$quality)
#str(smote_data1)
table(smote_data1$quality)
```

## SMOTE round 2 (X8)
```{r}
smote_data2 = SMOTE(select(smote_data1, -quality), select(smote_data1, quality), dup_size = 29)$data
smote_data2 = rename(smote_data2, quality = class)
smote_data2$quality = as.factor(smote_data2$quality)

table(smote_data2$quality)
```

## SMOTE round 3 (X4)
```{r}
smote_data3 = SMOTE(select(smote_data2, -quality), select(smote_data2, quality), dup_size = 13)$data
smote_data3 = rename(smote_data3, quality = class)
smote_data3$quality = as.factor(smote_data3$quality)

table(smote_data3$quality)
```

## SMOTE round 4 (X7)
```{r}
smote_data4 = SMOTE(select(smote_data3, -quality), select(smote_data3, quality), dup_size = 2)$data
smote_data4 = rename(smote_data4, quality = class)
smote_data4$quality = as.factor(smote_data4$quality)

table(smote_data4$quality)

data_bal = smote_data4
```


## Data Split
``` {r}
Y = make.names(data_bal$quality)
X = select(data_bal, -quality) 

trainSplit = createDataPartition(y = Y, p = 0.8, list = FALSE)

Ytrain = Y[trainSplit]
Xtrain = X[trainSplit,]
XtrainMat = as.matrix(Xtrain)

Ytest  = Y[-trainSplit]
Xtest  = X[-trainSplit,]
XtestMat = as.matrix(Xtest)
```

### Check for Class Imbalance
```{r}
table(Ytrain)
table(Ytest)

```



## Fit FDA (MARS) Classification Model
flexible discriminant analysis (FDA) using multivariate adaptive regression splines (MARS) basis functions

MARS is an algorithm that creates a piecewise linear model. It captures nonlinearity by automatically:
 * Allocating the bins ("knots")
 * Estimating the model in each bin 
Utilizes hinge functions 

Tuning parameters in MARS:
 * The degree of interactions between different hinge functions (degree)
 * the number/location of the knots (nprune)
 
 Use 5-fold cross validation 

> https://rdrr.io/cran/caret/man/bagFDA.html 

``` {r}
fdaOut = train(x = Xtrain, 
                y = Ytrain,
                method = 'fda',
                metric = 'Accuracy',
                tuneGrid = expand.grid(degree = 1:5, nprune = c(5,10,20,50,100)),
                trControl = trainControl(method='CV',number = 5,classProbs = TRUE))
fdaOut
plot(fdaOut)
```

## Best Model's CV Estimate of Testing Error 
```{r}
max(fdaOut$results$Accuracy)
```

## Feature Importance
``` {r}
fdaVip = vip(fdaOut,num_features = 11, bar = FALSE, metric = "Accuracy") + 
  ggtitle("Variable Importance Based on Accuracy") + 
  theme(plot.title = element_text(hjust = 0.5))
plot(fdaVip)
```

``` {r}
fdaOut$bestTune
coef(fdaOut$finalModel)
```



# Woah. 
```{r}
plot(fdaOut$finalModel)
```

## Prediction on Test Data
```{r}
YhatTestFDA = predict(fdaOut$finalModel, Xtest, type='class')
#YhatTestFDA    = ifelse(probHatTestFDA[,1] > 0.5, 'X1', 'X2')
```

The confusion matrices
```{r}
#table(YhatTestFDA, Ytest)
confusionMatrix(as.factor(YhatTestFDA), as.factor(Ytest))

#test accuracy
#cat("\nraw test accuracy:", mean(YhatTestFDA==Ytest))
```

Note: Model bad at predicting X3, X4 and X8 

## Multi-class ROC Curve 
```{r, warning=FALSE}
roc.multi <- multiclass.roc(Ytest, as.ordered(YhatTestFDA))
auc(roc.multi)
```


```{r}
rocs = roc.multi[['rocs']]
plot.roc(rocs[[1]])
sapply(2:length(rocs),function(i) lines.roc(rs[[i]],col=i))
```
